Computing in the Field of Rationals

### 2012-06-25 rational_ext.tex
_

The language {{Pi}} however lacks a natural representation for
first-class functions due to first-order nature of all of its
values. In the light of the work of duality of computation it was felt
that ``right model'' for expressing functions in {{Pi}} would also
express symmetries similar to those in the calculi of Filinski, Curien
and Herbelin and Wadler. The technical goal of this paper is to
present such a symmetric dual extension in the setting of {{Pi}}.

@@

\item Composition TODO.

@@

\noindent 
\textit{Example.}  We use the type {{bool}} as a shorthand to denote
the type {{1+1}} and use {{left ()}} to be {{true}} and {{right ()}}
to be {{false}}. The following combinator is represented by the given
diagram:

{{c : b * bool <-> b + b}}

{{c = swap* (;) dist (;) (identl* (+) identl*)}}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/example1-crop.pdf]
\includegraphics{diagrams/thesis/example1.pdf}
}
\end{center}

@@

\begin{proposition}[Strong Normalizing]
\label{prop:termination-pi} 
{{Pi}} computations always terminate.

{{forall. c:b1<->b2, v:b1, exists v':b2.}}  
{{c v |--> v'}}
\end{proposition}

@@

@@

For the purpose of establishing equivalence of morphisms in the
categorical presentation we use the following notion of equality:

\begin{lemma}[{{c1  = c2}}]
We say {{c1 = c2}} if {{c1:b1<->b2}} and {{c2:b1<->b2}} and for all
{{v1:b}}, we have {{c1 v1 |--> v2}} iff {{c2 v1 |--> v2}} and {{v2:b2}}.
\end{lemma}

>>>>

The language {{Pi}} however lacks a natural representation for
first-class functions due to the first-order nature of its
values. In the light of the work of duality of computation it was felt
that ``right model'' for expressing functions in {{Pi}} would also
express symmetries similar to those in the calculi of Filinski, Curien
and Herbelin and Wadler. The technical goal of this paper is to
present such a symmetric dual extension in the setting of {{Pi}}.

@@

\item Combinators can be composed is series ({{c1 (;) c2}}) or
  parallel. There are two forms of parallel composition -- combinators
  can be combined additively {{c1(+)c2}} (shown on the left) or
  multiplicatively {{c1(*)c2}} (shown on the right).

@@

%% \noindent 
%% \textit{Example.}  We use the type {{bool}} as a shorthand to denote
%% the type {{1+1}} and use {{left ()}} to be {{true}} and {{right ()}}
%% to be {{false}}. The following combinator is represented by the given
%% diagram:

%% {{c : b * bool <-> b + b}}

%% {{c = swap* (;) dist (;) (identl* (+) identl*)}}

%% \begin{center}
%% \scalebox{0.95}{
%% %%subcode-line{pdfimage}[diagrams/thesis/example1-crop.pdf]
%% \includegraphics{diagrams/thesis/example1.pdf}
%% }
%% \end{center}

@@

\begin{proposition}[Strong Normalizing]
\label{prop:termination-pi} 
{{Pi}} computations always terminate.  
{{forall. c:b1<->b2, v:b1, exists v':b2.}}  
{{c v |--> v'}}
\end{proposition}

@@

\begin{definition}[{{c1  = c2}}]
We say {{c1 = c2}} if {{c1:b1<->b2}} and {{c2:b1<->b2}} and for all
{{v1:b}}, we have {{c1 v1 |--> v2}} iff {{c2 v1 |--> v2}} and {{v2:b2}}.
\end{definition}

@@
